### 2022-10-31 Child-Metrics-Geo.Rmd
many updates not tracked

```{r}
x <- as.integer({{x}})
current_geo <- loop_df$tomatch[x]
current_geo_title <- ifelse(b_name %in% c("LAMIRDs"),
                            paste(current_geo, sub("s","",b_name)),
                            current_geo)
this_geo <- getGeoSFForRegion(loop_df,current_geo) 
this_comb_score <- getCombSFForRegion(this_geo)
has_roads <- TRUE # our assumption
if (is.logical(this_comb_score)) {
  has_roads <- FALSE
}
this_comb_total_score <- 0

```

#### `r current_geo_title`

```{r, results = 'asis', eval=!has_roads}
cat("This area does not have known pedestrian facilities.")
```

```{r, eval=has_roads}
this_comb <- getCombMapForRegion(comb_map, this_geo)
write_csv(this_comb_score, paste0(this_dir, 'region_score_ref/',b_name, x,'.csv'))

this_comb_total_score <- getAggregateScore(this_comb_score)
rm(this_comb_score)
```

```{r}
# add the score to the score breakdowns
if (!exists(deparse(substitute(this_comb_total_score)))){
  this_comb_total_score <- 0
}
score_breakdowns$Score[score_breakdowns$Area == current_geo] <- this_comb_total_score
```

```{r, eval=has_roads & has_points}
  # see if this geo has a point (is it an area around a place, like a school)
this_point <- st_as_sf(b_pointdf) %>%
      filter(NAME == this_geo$tomatch)
```

```{r, results = 'asis', eval=has_roads}
cat("This area has pedestrian facilities identified in the data; they are shown on the map for the left side and the right side of the road.")
```

```{r map-{{i}}{{x}}-sides, eval=has_roads}
#| fig.width = 14,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96
#|

p1 <- getMapForRegionInCounty(this_geo, current_geo) +
    labs(title = paste("Pedestrian Facilities:",current_geo_title))
r <- getMapFacilitiesSideForRegion(this_geo, "right") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} 
l <- getMapFacilitiesSideForRegion(this_geo, "left") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)}

  library(patchwork)

  p1 + r + l +
    plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

  detach("package:patchwork", unload=TRUE)
```

The overall score for pedestrian facilities in `r current_geo_title` is **`r nice_comma(this_comb_total_score)`**, **`r getScoreInterpretation(this_comb_total_score)`.**

```{r map-{{i}}{{x}}-scores, eval=has_roads}
#| out.width = '100%',
#| dpi = 96

  p1 <- getMapForRegionInCounty(this_geo, current_geo) +
    labs(title = paste("Calculated Pedestrian Facility Score:",current_geo_title))
  p2 <- getMapScoreSummaryForRegion(this_geo, this_comb, current_geo) +
    labs(title = "") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)}

  library(patchwork)

  p1 + p2 +
    plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

  detach("package:patchwork", unload=TRUE)
```

```{r kbl-{{i}}{{x}}, eval=has_roads}

getKableScoreSummaryForMap(this_comb)
```