### 2022-10-31 Child-Metrics-Region.Rmd

```{r bdsetup-{{i}}}
# create intro text
i <- as.integer({{i}})
has_points <- !is.na(breakdowns$pointdf[i])

b_name <- breakdowns$Name[i]
if(has_points){ b_pointdf <- eval(as.symbol(breakdowns$pointdf[i])) }
b_polydf <- eval(as.symbol(breakdowns$polydf[i]))
b_items <- eval(as.symbol(breakdowns$items[i]))
b_idcol <- breakdowns$idcolpoly[i]

loop_df <- st_as_sf(eval(b_polydf)) %>% mutate(tomatch = eval(as.symbol(b_idcol)))
score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = b_items, Score = NA))
```

### `r b_name`

```{r bdtext-{{i}}}
intro_text <- paste0("Kitsap County has ",
                    ifelse(!has_points, # poly only 
                           paste(nrow(b_polydf), 
                                 b_name)
                           , # now the ones with points
                           paste0(nrow(b_pointdf), 
                                  " ",
                                  b_name, 
                                  "; however, only ",
                                  nrow(b_polydf),
                                  " are outside of city limits. Here they are shown with a 15 minute walking distance")
                           ),".")

title_text <- paste("Kitsap County",b_name)
subtitle_text <- ifelse(b_name %ni% c("Commissioner Districts","LAMIRDs"),
                        "only showing the unincorporated county","")
caption_text <- paste("Data source:",breakdowns$data_source[i])

```

`r intro_text`

```{r bdmap-{{i}}}
#| fig.width = 8,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96
#| 
kitsap +
  geom_sf(data = b_polydf, aes(fill = eval(as.symbol(b_idcol))), 
          lwd = 0, alpha = 0.6) +
  {if (has_points) geom_sf(data = b_pointdf, aes(color = 1), 
                           color = "black", show.legend = FALSE) } +
  theme_void() +
  labs(title = title_text,
       #subtitle = subtitle_text,
       caption = caption_text,
       fill = b_name)

```

```{r bdloop-{{i}}, include=FALSE, echo=FALSE}

expanded_child <- lapply(
  seq_along(1:length(b_items)),
    function(xrow) {
      knitr::knit_expand(file = "Child-Metrics-Geo.Rmd",
                                x = xrow )
    })

parsed_child <- knitr::knit_child(text = unlist(expanded_child))
```

`r parsed_child`